---
title: "Seghier simulation"
output: html_notebook
---

This is a notebook script, which means that if you run it, it will generate a html file that combines the text and the code.

Attempt to reproduce what is done in Seghier, M. L. (2018). Categorical laterality indices in fMRI: a parallel with classic similarity indices. Brain Structure and Function, 224(3), 1377-1383. 

Just checking that I understand the method. Essentially, starts by simulating all possible L and R combinations of voxels, from 0 to N, where N can be specified by user. I have used N = 200 in interests of time here.


```{r makeAllLIs, include=TRUE, echo=TRUE}
LIlist<-vector()
N= 200 #n voxels
for (i in 0:N){ #possible values for L hem
  for (j in 0:N){ #possible values for R hem
    LIlist<-c(LIlist,(i-j)/(i+j))
  }
}
#Seghier shows situation with many breaks, which gives quite a complex distribution. However, with fewer breaks it looks more normal.
hist(LIlist,breaks=5000)
hist(LIlist,breaks=100)

LIcode<-rep(2,length(LIlist))

T <- .2 #cutoff found to give equal Ns in 3 groups
LIcode[LIlist>-T]<-1#recode as 1 for those who are below -T
LIcode[LIlist>T]<-3 #recode as 3 for those who are above T
table(LIcode)

```

# LI and classic dissimilarity indices

This sounds complicated, but is really quite straightforward.
From the text: Given that LI distribution is symmetric, focus on positive LI values.

Rearrange suprathreshold voxels in a homologous way between left and right ROI, so each selected voxel on left has a homologue on the right.
Does not affect LI but allows voxel by voxel comparison.

Can then say LH-RH = OR(LH,RH) - AND(LH,RH) #OR here is logical operator, not odds ratio!
and 
LH+RH = OR(LH,RH)+AND(LH,RH)

so OR(LH,RH) is instance where L or R is active
AND(LH,RH) is instance where L and R are active

and because voxels are homologous
OR(LH,RH) = max(LH,RH)
AND(LH,RH) = min(LH,RH)

Easier to understand with a concrete example.

Suppose we have situation with 7 active voxels in the R, and 20 active voxels in the L. 
Then OR(LH,RH) = 20 (because all the active voxels in R have homologue in L)
AND(LH,RH)= 7  (ie voxels where both are active)

This fits with the min and max definitions above.

We can then relate this to the Sorensen-Dice (SD) index, which is

SD = 2AND/(OR+AND)

In our case this would be 14/27 = .5185

and LI is 1-SD = .4814 (which is same as (L-R)/(L+R))

So the LI can be related to classic similarity indices, if we assume maximal voxel-wise overlap after spatial rearrangement.
Can relate to rich literature on classic indices in taxonomy.

Sorensen-Dice index is special case of the kappa statistic that is used in reliability.
Strong agreement or similarity between quantitites is commonly considered to be above .8

This means that LI less than .2 corresponds to highly similar LH and RH.

1-SD is not a proper distance metric because it does not possess property of triangle inequality. (I have no idea what this is!)
But we can redefine it to convert LI to a proper distance.
Here used Jaccard-Tanimoto index, which is a proper distance. (Again, I am taking this on trust!)

JT is defined as AND/OR
and LInew is 1-JT

then the new LI is (LH-RH)/max(LH,RH)

Works to give equal distributions with cutoff T of 1/3

Can simulate this new LI just as before. This agrees with the account in the paper.

```{r makeLInew,include=TRUE,echo=TRUE}
LInew<-vector()
N= 200 #n voxels
for (i in 0:N){ #possible values for L hem
  for (j in 0:N){ #possible values for R hem
    LInew<-c(LInew,(i-j)/max(i,j))
  }
}
hist(LInew,breaks=5000)
hist(LInew,breaks=100)
LIcode<-rep(1,length(LInew))
T <- 1/3 #cutoff found to give equal Ns in 3 groups
LIcode[LInew>-T]<-2
LIcode[LInew>T]<-3
table(LIcode)

```
Can also convert LI to LInew with (2*LI)/(1+abs(LI))
